### 2020-02-03 nlp_models.tex
updating nlp

However, as seen in Figure \ref{fig:GPT-snorm-hist},
the ``bad'' GPT model has, spuriously, many smaller Spectral Norms $\log\Vert\mathbf{W}\Vert_{\infty}$,
compared to the ``good'' GPT2--which violates the belief that smaller Spectral Norms are always better.
Indeed, because there are so many anonymously small $\Vert\mathbf{W}\Vert_{\infty}$,
it appears that the GPT model may be exhibiting a kind of soft rank collapse and/or undesired
loss of Frobeinus mass.
This is an extremely important observation because it demonstrates that while the Spectral Norm
may correlate well with predicted test error, it is not a good indicator of the overall quality of a model,
and using it as an empirical metric may give spurious results when applied to poorly trained
or otherwise deficient models.

@@

also differs significantly between GPT and GPT2.
Figure \ref{fig:gpt-alpha-layer} plots $\alpha$ vs the depth (i.e. a layer id) for each model.

>>>>

However, as seen in Figure \ref{fig:GPT-snorm-hist},
the ``bad'' GPT model has, spuriously, many smaller Spectral Norms $\log\Vert\mathbf{W}\Vert_{\infty}$,
compared to the ``good'' GPT2--which violates the belief that smaller Spectral Norms are always better.
Indeed, because there are so many anonymously small $\Vert\mathbf{W}\Vert_{\infty}$,
it appears that the GPT model may be exhibiting a kind of \emph{scale collapse} and/or undesired
loss of Frobeinus mass.
This is an extremely important observation because it demonstrates that while the Spectral Norm
may correlate well with predicted test error, it is not a good indicator of the overall quality of a model,
and using it as an empirical metric may give spurious results when applied to poorly trained
or otherwise deficient models.

@@

also differs significantly between GPT and GPT2.
Figure \ref{fig:gpt-alpha-layer} plots $\alpha$ vs the depth (i.e. a layer id) for each model.
The deficient GPT model displays two trends in $\alpha$ , one stable with $\alpha\sim 4$,
and one increasing with layer id, with $\alpha$ reaching as high as $12$.
In contrast, the well-trained GPT2 model shows consistant and stable patterns, again
with one stable $\alpha\sim 3.5$ (and below the GPT trend), and the other only
slightly trending up, with $alpha\le 6$. 
Again, the scale-invariant $\alpha$ metric lets us identify potentially
poorly-trained models.

Notice also that the Spectral Norms
(Figure \ref{fig:gpt-snorm-layer})
and the $\alpha-$Norms
(Figure \ref{fig:gpt-pnorm-layer})
display an unusually some mall values for GPT,
and some increasing trending values for GPT2.
Generally speaking, we do not think scale-dependennt 
norm metrics can be directly applied to distinguish ``good'' vs ``bad''  models 
because of the anamalous scale collapse that is observed in bad models.
